### 2021-06-02 PartIII.tex
updating by O

\documentclass[11pt]{amsart}

\usepackage{macros, setspace}

\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}

%trying this out
%\usepackage[upint]{stix}
%\usepackage{cmupint}

\def\brian#1{{\textcolor{blue!65!red}{BRW: {#1}}}}
\def\owen#1{{\textcolor{green!65!black}{OGG: {#1}}}}

\author{Owen Gwilliam and Brian R. Williams}
\date{\today}

\title{Part III}

\spacing{1.25}

\begin{document}
\maketitle

\section{Vertex algebras and higher-dimensional analogs \owen{Crappy title}}

So how do factorization algebras appear in holomorphic field theories?

\subsection{Chiral conformal field theories and vertex algebras}

\owen{I want to discuss how VOAs relate to factorization algebras, notably our CDO project, Si's results, and Brueggeman's thesis. We should also gesture at the Beilinson-Drinfeld approach.}

\subsection{Higher Kac-Moody algebras and their free field realizations}

\subsection{Holomorphic $\sigma$-models and higher-dimensional chiral differential operators}

\subsection{Work with Ingmar \owen{not the title}}

\end{document}