### 2023-08-22 highervertexops.tex
alternatives for 2d algebra

\documentclass[11pt]{amsart}

\usepackage{macros, setspace}

\spacing{1.1}

\usepackage{parskip}
\setlength{\parindent}{18pt}
\setlength{\parindent}{0cm}

\author{Owen Gwilliam and Brian R. Williams}
\date{\today}
\title{Holomorphic field theories and higher algebra}

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}
\def\del{\partial}
\def\til{\Tilde}
\def\Sol{\text{Sol}}
\def\Spec{\text{Spec}}
\def\C{{\mathbb{C}}}
\def\RHom{{\rm RHom}}
\newcommand{\sff}{\mathsf{f}}
\newcommand{\sfe}{\mathsf{e}}
%\def\sfR{\mathsf{R}}
%\def\sfA{\mathsf{A}}
\newcommand{\laurent}{\Hat \sfA}
\newcommand{\sfb}{\mathsf{b}}
\newcommand{\sfc}{\mathsf{c}}
\DeclareMathOperator{\colim}{colim}

\renewcommand{\op}{\operatorname}
\def\lie#1{\ensuremath{\mathfrak{#1}}}

\def\brian#1{{\textcolor{blue!65!red}{BW: {#1}}}}
\def\owen#1{{\textcolor{violet!65!black}{OG: {#1}}}}

%\usepackage[upint]{stix}

\begin{document}

\section{Overview of definitions and goals}

The goal here is to record higher-dimensional analogs of the ingredients for defining a {\em field} in a holomorphic theory (i.e., a higher-dimensional {\em vertex operator}). We do not aspire (yet!) to give the higher-dimensional version of a vertex algebra.

{\it Notations for spaces, functions, and distributions}

Let $\cO_d$ denote a (preferably ``small'' or convenient) model of the derived sections of the structure sheaf (of {\em algebraic} functions) on the punctured affine space $\mathring{\AA}^d$.\\
{\tiny When $d=1$, it is the Laurent polynomials~$\CC[z,z^{-1}]$.}

Let $\widehat{\cO}_d$ denote a (preferably ``small'' or convenient) model of the derived sections of the structure sheaf (of {\em algebraic} functions) on the punctured formal disk $\mathring{\DD}^d$.\\
{\tiny When $d=1$, it is the Laurent series $\CC((z))$.}

Let $\cD_d$ denote a model of the derived ``distributions'' on the punctured affine space $\mathring{\AA}^d$. It is the linear dual to~$\cO_d$.\\
{\tiny When $d=1$, it is $\CC[[z,z^{-1}]]$, which is a notation that indicates infinite formal sums in both directions. The pairing is via the residue pairing. (This notation is poor because it suggests an algebra, but it is not.)}

Let $\widehat{\cD}_d$ denote a model of the derived ``distributions'' on the punctured formal disk $\mathring{\DD}^d$. It is the linear dual to~$\widehat{\cO}_d$.\\
{\tiny When $d=1$, it is again $\CC((z))$, via the residue pairing.}

There are analogs of these constructions for products of affine or formal spaces with the fat diagonal removed.
We will write here the two-variable version.

Consider the space 
\[
\mathring{\AA}^{d}[2]= (\AA^d \times \AA^d) - \Delta_f{[2]}, 
\]
where $\Delta[2]_f$ denotes the fat diagonal
\[
\AA^d \times \{0\} \cup \{0\} \times \AA^d \cup \AA^{2d}_\Delta
\]
with
\[
 \AA^{2d}_\Delta = \{ (x,y) \in \AA^d \times \AA^d \; : \; x \neq y\}.
\]
Let $\mathring{\DD}^d[2]$ denote the corresponding formal space.

Let $\cO_d{[2]}$ denote a model of the derived sections of the structure sheaf (of {\em algebraic} functions) on $\mathring{\AA}^d[2]$.

Let $\widehat{\cO}_d{[2]}$ denote a model of the derived sections of the structure sheaf (of {\em algebraic} functions) on $\mathring{\DD}^d[2]$.

Let $\cD_d[2]$ denote a model of the derived distributions on $\mathring{\AA}^d[2]$.

Let $\widehat{\cD}_d[2]$ denote a model of the derived distributions on $\mathring{\DD}^d[2]$.
\newpage

{\it Operators}

Let $V$ be a cochain complex, which we will treat as a {\em state space}.

A {\em field} is an element of 
\[
\RHom(V \otimes \widehat{\cO}_d, V) \cong \RHom(V, V \widehat{\otimes} \widehat{\cD}_d).
\]
When $d=1$, this definition becomes the usual version: $\Hom(V, V((z)))$.

Note that a field can be seen as a distribution that eats a function on the formal disk and gives a linear operator on $V$ because
\[
\RHom(V \otimes \widehat{\cO}_d, V) \cong \RHom(\widehat{\cO}_d, \RHom(V,V)).
\]
This corresponds to a traditional notion of a ``quantum field'' as an operator-valued distribution.

For notation, if $z$ denotes a coordinate $(z_1,\ldots,z_d)$ for the formal disk $\DD^d$,
we write $a(z)$ for a field using that coordinate.

Given two fields $a(z)$ and $b(w)$, we have a commutator $[a(z),b(w)]$ that eats a function in $\widehat{\cO}_d[2]$ and returns an endomorphism of~$V$.

Two fields are {\em mutually local} if there exists a natural number $N \in \NN$ and an element $f(z,w)$ of the ideal $(z_i-w_i)_{i =1}^d \subset \widehat{\cO}(\DD^d\times \DD^d)$ such that $f(z,w)[a(z),b(w)]$ is cohomologous to zero (i.e., it is exact).

This condition recovers the usual notion of locality after passing to cohomology. 
\owen{I suspect there's something better to say here, informed by your work with Nik. I'm least happy with this definition.}

It is possible to formulate a representative of the ``delta function'' and one can hope that the usual formulas (e.g., describing the structure of the singular parts of commutators or OPEs) have analogs up to exact terms.

{\it Free field realizations}

We want to show that the $L_\infty$-action of a higher Kac-Moody algebra on the state space of a free theory is via mutually local fields.

\newpage

\section{Models for functions and distributions on punctured affine space}

We want a convenient and explicit model for the commutative algebra-up-to-homotopy $\RR \Gamma(\mathring{\AA}^2)$.
Our approach follows the approach of Keyou Zeng but is written in a more algebraic style.
We quickly recall his results before introducing our version.

\subsection{Zeng's hyperplane approach}

The punctured affine space $\mathring{\AA}^2$ contains a natural real hypersurface given by the unit sphere $S^3 = \{|w_1|^2 + |w_2|^2 = 1\}$, with the $w_i$ being holomorphic coordinates on $\AA^2$.
The theory of CR manifolds provides an analogue of the Dolbeault complex, 
and Zeng works with polynomial functions and polynomial forms within it:\footnote{Zeng claims to compute the CR cohomology but he in fact just works with the direct sum of the $SU(2)$-representations therein, and hence he works with this algebraic subcomplex instead.}
\beqn
\Omega^{0,\bu}_b (S^3) = \big(\C[w_i,\Bar{w}_i, \xi_i] \slash (w_i \Bar{w}_i = 1, w_i \xi_i = 0) \big) .
\eeqn
where the $\xi_i$ have degree 1 (and represent $\d \Bar{w}_i$) and where the differential is $\dbar_b = \xi_i \del_{\Bar{w}_i}$.
Note that one can see that polynomials in the coordinates, both holomorphic $w_i$ and antiholomorphic $\Bar{w}_i$, span the complex-valued polynomial functions on $\CC^2 \cong \RR^4$, i.e., as a real vector space.
Their restriction to the unit-sphere $S^3$ wholly determines their behavior as functions on $\RR^4$, and hence also on punctured affine space. 
The elements $\xi_i$ correspond to $d\Bar{w}_i$ in the Dolbeault complex of punctured affine space.
The differential $\dbar_b$ then imposes the constraint, in degree zero, that these functions be holomorphic,
and its image is the image of $\dbar$ intersected with polynomial  forms in $d \Bar{w}_i$.
Thus the cohomology of this complex $\Omega^{0,\bu}_b (S^3)$ is a model for $\RR \Gamma(\mathring{\AA}^2)$, by a result of~\cite{FHK}.

Zeng analyzes this complex using the natural $SU(2)$-action on the 3-sphere, which is the restriction of the fundamental representation to this hypersurface.
More accurately, $SU(2)$ acts on the linear span of $w_1$ and $w_2$ by the fundamental representation and on the linear span of $\Bar{w}_1$ and $\Bar{w}_2$ by the conjugate (or anti-fundamental) representation, 
and it respects the relation $\sum_i w_i \Bar{w}_i = 1$ and the differential $\dbar_b$, 
so the action descends to the whole complex.
One can decompose the complex into irreducible representations, 
allowing one to identify the cohomology as a subcomplex.
This construction yields an explicit deformation retraction of the $\dbar_b$-complex onto its cohomology.

\subsection{An algebraic version}

We mimic the construction but replace the role of $SU(2)$ by $SL(2)$.
\owen{I mostly left things alone below, except for removing typos or modestly rewording things I found confusing.}

Let $\sfR$ be the commutative algebra generated by variables $z_i,\lambda_i,i=1,2$ subject to the relation $z_1\lambda_1 + z_2 \lambda_2 = 1$.
Let $\sfA$ be the graded commutative algebra freely generated over $\sfR$ by a degree $+1$ element $\omega$.
Thus $\sfA = \sfR \oplus \sfR \omega [-1]$.
Equip it with the differential
\[
\dbar(z_i) = 0 \quad\text{and}\quad \dbar(\lambda_i) = \epsilon_{ij} z_j \omega
\]
extended as a derivation.
That is, $\dbar(\lambda_1) = z_2 \omega$ and $\dbar(\lambda_1) = -z_1 \omega$.

We view the $z_i$ as the $w_i$ in Zeng's complex, and the $\lambda_i$ as the $\Bar{w}_i$.
The element $\omega$ corresponds to $\Bar{w}_1\xi_2 - \Bar{w}_2\xi_1$.
In fact, the key result of this section is the following.

\begin{prop}
There is a $SU(2)$-equivariant isomorphism of dg commutative algebras
\beqn
\Phi \colon \sfA \to \Omega^{0,\bu}_b (S^3) 
\eeqn
defined by $\Phi(z_i) = w_i$, $\Phi(\lambda_i) = \Bar{w}_i$, and $\Phi (\omega) = \ep_{ij} \Bar{w}_i \xi_j$.
\end{prop}

This proposition guarantees that we have a well-behaved model for~$\RR \Gamma(\mathring{\AA}^2)$.

The algebra $\sfR$ is naturally an $SL(2)$ representation where the linear span $V_z$ of $z_1$, $z_2$ transforms in the fundamental representation  and the linear span $V_\lambda$ of  $\lambda_1$, $\lambda_2$ transforms in the dual (or contragredient) representation.
In particular, with respect to the standard Cartan subgroup, $z_1$ and $\lambda_2$ have weight 1 and $z_2$ and $\lambda_1$ have weight~$-1$.

Consider now the space of all polynomials
\[
\C[z_i,\lambda_i] = \Sym(V_z \oplus V_\lambda) \cong \Sym(V_z) \otimes \Sym(V_\lambda).
\]
The Lie algebra $\sl(2)$ acts by the lowering and raising operators
\beqn
\sff = z_2 \del_{z_1} - \lambda_1 \del_{\lambda_2} \quad\text{and}\quad \sfe = z_1 \del_{z_2} - \lambda_2 \del_{\lambda_1}.
\eeqn
The space of polynomials breaks up into a direct sum over the subrepresentations 
\[
V_{a,b} = \Sym^a(V_z) \otimes \Sym^b(V_\lambda)
\]
of polynomials that are homogenous of order $a$ in the variables $z_i$ and homogenous of order $b$ in the variables~$\lambda_i$.

Recall that $\Sym^a(V_z)$ is an irreducible representation of dimension $a+1$ (and hence highest weight $a$) and $\Sym^b(V_\lambda)$ is an irreducible representation of dimension $b+1$ (and hence highest weight $a$).
Let $I_m$ denote the irreducible representation of highest weight $m$ (and hence of dimension $m+1$).
By standard facts about $SL(2)$, we see that $V_{a,b}$ decomposes as a direct sum of irreducible representations
\begin{equation}
\label{eq:sl2decompV}
V_{a,b} \cong I_{a+b} \oplus I_{a+b-2} \oplus \cdots \oplus I_{a-b}
\end{equation}
for $a \geq b$. (The other situation can be seen by symmetry.)

There is a more concrete way to describe this decomposition, 
modeled on the construction of spherical harmonics as representations of $SO(n)$.
Consider the element  
\[
|(z,\lambda)|^2 \define z_1 \lambda_1 + z_2 \lambda_2, 
\]
which describes the evaluation pairing. 
It is fixed by the $SL(2)$-action, has weight zero, and bigrading $(1,1)$.
Multiplication by this element produces an injection
\[
V_{a-1,b-1} \hookrightarrow V_{a,b}
\]
that is $SL(2)$-equivariant, so we know there is a decomposition 
\[
V_{a,b} \cong |(z,\lambda)|^2 V_{a-1,b-1} \oplus C_{a,b},
\]
where $C_{a,b}$ is a complementary subspace and representation.
We can characterize this subrepresentation concretely using the differential operator
\beqn
\triangle = \del_{z_1} \del_{\lambda_1} + \del_{z_2} \del_{\lambda_2}
\eeqn
acting on $\C[z_i,\lambda_i]$;
it is inspired by the Laplacian on $\CC^2$ when we view $\lambda_i$ as a version of $\Bar{z_i}$.
This operator has weight zero and, in fact, it is $SL(2)$-equivariant.\footnote{For example, notice that for any $g(z_i,\lambda_i)$ we have
\begin{align*}
\triangle \circ \sff(g) &= (\del_{z_1} \del_{\lambda_1} + \del_{z_2} \del_{\lambda_2})(z_2 \del_{z_1} g - \lambda_1 \del_{\lambda_2}g )\\
&= z_2 \del_{z_1}^2 \del_{\lambda_1} g - \del_{z_1} \del_{\lambda_2} g - \lambda_1 \del_{z_1} \del_{\lambda_1} \del_{\lambda_2} g \\ & \qquad + \del_{z_1} \del_{\lambda_2}  g + z_2 \del_{z_1} \del_{z_2} \del_{\lambda_2} g - \lambda_1 \del_{z_2} \del_{\lambda_2}^2 g \\ 
& = z_2 \del_{z_1}^2 \del_{\lambda_1} g - \lambda_1 \del_{z_1} \del_{\lambda_1} \del_{\lambda_2} g + z_2 \del_{z_1} \del_{z_2} \del_{\lambda_2} g - \lambda_1 \del_{z_2} \del_{\lambda_2}^2 g \\
&=  \sff \circ \triangle(g)
\end{align*}
Thus $[\triangle,\sff] = 0$.
Similarly, one can show that $[\triangle, \sfe] = 0$.)
\owen{I need to check this myself. If there's a slick way, we should include it.} \brian{I gave a computational proof.}\owen{Thank you! I don't know why I'm so lazy about calculus.}}
On the other hand, it maps $V_{a,b}$ to $V_{a-1,b-1}$, i.e., it changes the bigrading by~$(-1,-1)$.

Let $H_{a,b} = \ker \triangle|_{V_{a,b}}$ denote the ``harmonic'' polynomials of bidegree $(a,b)$.
Examples of harmonic polynomials include $z_1^a \lambda_2^b$ for arbitrary $a,b$ and
$z_1 \lambda_1 - z_2 \lambda_2$ for $a = 1 = b$.
Direct computation shows that 
\[
V_{a,b}  = H_{a,b} \oplus |(z,\lambda)|^2 V_{a-1,b-1},
\]
and by iteratively applying this decomposition, we find
there is a direct sum decomposition
\[
V_{a,b}  = H_{a,b} \oplus |(z,\lambda)|^2 H_{a-1,b-1} \oplus |(z,\lambda)|^4 H_{a-2,b-2} \oplus \cdots \oplus |(z,\lambda)|^{2b} H_{a-b,0} 
\]
when $a \geq b$.
Thus $C_{a,b} = H_{a,b}$, and we also find $I_{a+b} = H_{a,b}$ by comparing with the decomposition~\eqref{eq:sl2decompV}.

In light of these facts, we see that the canonical quotient map 
\[
q \colon \C[z_i,\lambda_i] \to \sfR = \C[z_i,\lambda_i]/(|(z,\lambda)|^2 = 1)
\]
is $SL(2)$-equivariant.
Moreover, inside $V_{a,b}$, any multiple of $|(z,\lambda)|^2$ is annihilated by $q$,
so $\sfR$ is spanned by the images of $H_{a,b}$.
Thus, there is a direct sum decomposition 
\beqn
\sfR = \bigoplus_{a,b} \sfR_{a,b}
\eeqn
where $\sfR_{a,b} = q(H_{a,b})$ is the image of the subspace of harmonic polynomials in $\C[z_i,\lambda_i]$ under the quotient map.

We emphasize that this decomposition does \textit{not} respect the commutative product.
\owen{I'm confused what you mean here. Can you spell out what you mean in more detail?}

\begin{eg}
The space $\sfR_{a,0}$ consists of polynomials in $z_1,z_2$ of homogenous degree~$a$.
The space $\sfR_{0,b}$ consists of polynomials in $\lambda_1,\lambda_2$ of homogenous degree $b$.
The space $\sfR_{1,1}$ is three-dimensional with a basis given by
\beqn
z_1 \lambda_2, z_1 \lambda_1 - z_2 \lambda_2, z_2 \lambda_1 .
\eeqn
\end{eg}

The following fact is quite useful for us.

\begin{lem}
Each $\sfR_{a,b}$ is isomorphic to the irreducible $SL(2)$-representation of dimension $(a+b+1)$.
Its highest weight vector is $z_1^a \lambda_2^b \in \sfR_{a,b}$, and a basis for $\sfR_{a,b}$ is obtained as
\beqn
\{z_1^a \lambda_2^b, \sff (z_1^a \lambda_2^b), \sff^2 (z_1^a \lambda_2^b), \ldots, \sff^{a+b} (z_1^a \lambda_2^b) \} .
\eeqn
\end{lem}

This lemma is proved by applying the description of the decomposition into irreducibles.
In turn, any polynomial $f(z,\lambda) \in \sfR$ admits a decomposition of the form
\beqn\label{eqn:expand}
f(z, \lambda) = \sum_{a,b,k \geq 0} c_{k}^{ab} \sff^k (z_1^a \lambda_2^b) 
\eeqn
for some coefficients $c_k^{ab}$.

By direct inspection, one finds that the $\dbar$-operator is $SL(2)$-equivariant.

\begin{cor}
The differential $\dbar$ provides an isomorphism 
\[
\dbar_{a,b} \colon \sfR_{a,b} \xto{\cong} \sfR_{a+1,b-1} \omega
\]
for $b > 0$.
Thus, the cohomology of $\sfA$ is 
\[
H^0(\sfA) = \bigoplus_{a \in \NN} \sfR_{a,0}
\]
and
\[
H^1(\sfA) = \bigoplus_{b \in \NN} \sfR_{0,b}.
\]
\end{cor}

\begin{proof}
For the first claim, note that the map is equivariant and nontrivial when $b > 0$. Thus by Schur's lemma it is an isomorphism.
The description of cohomology follows by keeping the summands $\sfR_{a,b}$ that appear in the kernel (for $H^0$) and that do not appear in the image (for $H^1$).
\end{proof}

%Define the involution $\Bar{(-)}$ on $\sfR$ by $\zbar_i = \lambda_i$ and consider the inner product on $\sfR$ given by the formula
%\beqn
%\<a,b\> = \op{Res} \left(\Bar{a} b \omega \right).
%\eeqn
%This inner product is $SU(2)$ invariant.
%
%\begin{eg}
%An orthonormal basis for $\sfR_{a,0}$ is
%\beqn
%e_m^{(a,0)} = \sqrt{\frac{(a+1)!}{m! (a-m)!}} z_1^m z_2^{a-m} , \quad m = 0,1,\ldots,a .
%\eeqn
%Likewise, an orthonormal basis for $\sfR_{0,b}$ is
%\beqn
%e_m^{(0,b)} = \sqrt{\frac{(a+1)!}{m! (a-m)!}} (-\lambda_1)^m \lambda_2^{a-m} , \quad m = 0,1,\ldots,b .
%\eeqn
%An orthonormal basis for $\sfR_{1,1}$ is
%\beqn
%e_2^{(1,1)} = \sqrt{6} z_1 \lambda_2, \quad e_1^{(1,1)} = \sqrt{3}(z_1 \lambda_1 - z_2 \lambda_2), \quad e_0^{(1,1)} = \sqrt{6} z_2 \lambda_1 .
%\eeqn
%\end{eg}

\begin{proof}[Proof of proposition]
In fact $\Phi$ maps the subspace $\sfR_{a,b}$ isomorphically to $\cH_{a,b}$ where $\cH_{a,b} \subset \Omega^{0,\bu}_b (S^3)$ is the space of harmonic polynomials that are of bidegree $(a,b)$ in the variables $(w_i,\Bar{w}_j)$.
Now observe that
\beqn
\Phi(\dbar(\lambda_i)) = - \Phi(\ep_{ij} z_j \omega) = - \ep_{ij} w_j \Phi (\omega) = \xi_i = \dbar_b \Phi (\lambda_i) 
\eeqn
to see that the differentials are compatible.
\end{proof}

\subsection{Homotopy data}

Let $\sfH \define H^\bu(\sfA)$ and let $i,p$ be the $SU(2)$-equivariant inclusions and projections $i \colon \sfH \hookrightarrow \sfA$, $p \colon \sfA \to \sfH$, given by the direct sum description of the corollary above.

There is a canonical contracting homotopy $H$ on $\sfA$ 
because for each $b > 0$, there is a canonical equivariant inverse map $\dbar_{a,b}^{-1}:  \sfR_{a+1,b-1} \omega \to \sfR_{a,b}$.
Explicitly, send the highest weight vector $z_1^{a+1} \lambda_2^{b-1} \omega$ to the highest weight vector $z_1^{a} \lambda_2^{b}$ and then determine where other vector go via the lowering operator.
For $b = 0$, $H|_{\sfR_{0,b}\omega} = 0$.

\owen{Check the side conditions!}

\begin{lem}
The data $(i,p,H)$ determines a special deformation retraction where $i$ is a map of dg commutative algebras.
\end{lem}

Note that $p$ is not a map of algebras. 
For example,
the following three polynomials in $\C[z_i,\lambda_i]$ represent the same element in~$\sfR$:
\beqn
\frac12  + \frac12 (z_1 \lambda_1 - z_2 \lambda_2) \sim \frac12  + \frac12 (z_1 \lambda_1 - z_2 \lambda_2) + \frac12 (z_1 \lambda_1 + z_2 \lambda_2) \sim z_1 \lambda_1
\eeqn
so
\beqn
p(z_1 \lambda_1) = p\left(\frac12  + \frac12 (z_1 \lambda_1 - z_2 \lambda_2)\right) = \frac12.
\eeqn

\subsection{Functions on punctured formal disk}

The model $\sfA$ for algebraic functions on punctured affine space was constructed using the ring
\beqn
\sfR = \C[z_i,\lambda_i] \slash (\sum_i z_i \lambda_i = 1) .
\eeqn
Let $\Hat{\sfR}$ denote the completion of this ring with respect to the ideal generated by~$z_1,z_2$:
\beqn
\Hat{\sfR} = \C[[z_i]][\lambda_i] \slash (\sum_i z_i \lambda_i) .
\eeqn
\owen{As you said originally, this second line takes the completion for the $z$s first and then adjoins the $\lambda$s, while your definition involves completing after adjoining the $\lambda$s. It's not clear those operations commute.}

Let $\Hat{\sfA}$ be the graded algebra free generated over $\Hat{\sfR}$ and the degree one variable $\omega$.
The cohomology of $(\Hat\sfA, \dbar)$ is $\C[[z_i]]$ in degree zero and $\C[\lambda_i]$ in degree one.
\owen{We need to be really careful about this, and it probably depends on what category of vector spaces (e.g., complete filtered?) we work in. The $SL(2)$-approach above manifestly breaks down, although it gives us useful information.}

\subsection{The higher algebraic residue}

Define
\beqn
\op{Res} \colon \sfA \to \C[-1]
\eeqn
by the rules that $\op{Res}|_{\sfA^0} = 0$ and
\beqn
\op{Res} \left(f(z,\lambda) \omega \right) = p (f(z,\lambda)) (z=0) .
\eeqn
In terms of the expansion \eqref{eqn:expand} we have
\beqn
\op{Res} \left(f(z,\lambda) \omega \right) = c_{0}^{00} .
\eeqn

This residue extends to the completion to give a map
\beqn
\op{Res} \colon \laurent \to \C[-1] .
\eeqn

\begin{lem}
\label{lem:res} 
The following are true regarding the higher algebraic residue:
\begin{enumerate}
\item The residue is $\lie{sl}(2)$-invariant.
\item For $f \in \C[[z_1,z_2]]$ one has $\op{Res}(f \omega) = f(0)$.
\item For any $f \in \laurent$ one has $\op{Res}(\dbar f) = 0$.
Thus $\op{Res}$ is a cochain map.
\end{enumerate}
\end{lem}
\begin{proof}
Item (1) holds since $p$ is the $\lie{sl}(2)$-equivariant projection. 
For (2), note that for $f \in \C[z_1,z_2]$ that $p(f) = f$.

Finally, observe that for $f = z_1^a \lambda_2^b$ we have
\beqn
\op{Res}(\dbar f) = \op{Res}(b z_1^{a+1} \lambda_2^{b-1}) .
\eeqn
For $b =0$ this expression clearly vanishes. 
For $b = 1$ the right hand side is $p( z_1^{a+1})(0) = 0$ for all $a \geq 0$.
Finally, for $b > 1$ one has $p(z_1^{a+1} \lambda_2^{b-1}) = 0$ for all $a \geq 0$.
Using $\lie{sl}(2)$ invariance this proves (3).
\end{proof}

In the single-variable case, the residue map defines an identification
\beqn
\C((z)) \simeq \Omega^1 (\Hat{D})^\vee = \left(\C((z)) \d z\right)^\vee
\eeqn
where the right hand side is the continuous linear dual.
Explicitly, this identification takes a Laurent series $f$ to the functional 
\beqn
g \, \d z \mapsto \op{Res}_{1d} (f g \d z) .
\eeqn
In this way, distributions and functions on the punctured formal disk are isomorphic.

\begin{prop}\label{prop:res}
The higher algebraic residue defines a cochain map
\beqn
\sfA \to \left(\sfA \, \d^2 z\right)^\vee [-1]
\eeqn 
that sends an element $a \in \sfA$ to the functional $F_a$ where 
\beqn
F_a(b \, \d^2 z) =  (-1)^{|a|} \Res(a b \, \d^2 z) .
\eeqn
\brian{weird sign, but I think it's there because of the $[-1]$ shift.}
When extended to the completed version of punctured affine space, this assignment defines an isomorphism of cochain complexes 
\beqn
F \colon \Hat\sfA \simeq \left(\Hat\sfA \, \d^2 z\right)^\vee [-1] .
\eeqn
\end{prop}

\begin{proof}
Suppose that $a \in \sfA^0$. 
Then
\begin{align*}
F_{\dbar a} (b\, \d^2 z) & = - \Res \left( (\dbar a) b \,\d^2 z \right) \\
& = - \Res\left( \dbar(a b \,\d^2 z) \right) + \Res \left(a \dbar b \,\d^2 z\right) \\
& = F_{a} (\dbar b \,\d^2 z) .
\end{align*}
In the final line we have used (3) from lemma \ref{lem:res}.
If $a \in \sfA^1$ then $F_{\dbar a}(b \,\d^2 z) = 0 = F_{a} (\dbar b \,\d^2 z)$ for all $b$.

\brian{Can you help me prove part 2?}
\end{proof}

\section{Algebraic free field realization}

Consider the vector space $V = \laurent$.
Via the residue pairing we have an identification $V^\vee = \laurent \, \d^2 z [1]$ (see proposition \ref{prop:res}).
From hereon we use the notation $\laurent^{(2)} = \laurent \, \d^2 z$.
We will denote linear elements in the corresponding Poisson algebra
\beqn
\cP = \Sym \left(\laurent  \oplus \laurent^{(2)}[1] \right) ,
\eeqn
by $(\sfb , \sfc)$.
On these elements, the Poisson bracket reads
\beqn
\{\sfb ,\sfc\} = \Res(\sfb \sfc) .
\eeqn
The Lie algebra $\cP_2$ is spanned by symbols of the form $\sfb \sfb',\sfb \sfc, \sfc \sfc'$ where $\sfb,\sfb'\in \laurent$ and $\sfc,\sfc' \in \laurent^{(2)}[1]$.

The algebra $\cP$ is the free commutative algebra generated by $\laurent \oplus \laurent^{(2)}[1]$.
We can view, via the residue pairing, a linear element $(\sfb, \sfc) \in \laurent \oplus \laurent^{(2)}[1]$ as the linear functional
\[
\begin{array}{cccc}
(\sfb, \sfc) \colon & \laurent^{(2)} [1] \oplus \laurent & \to & \C\\
& (\beta, \gamma) & \mapsto  &(-1)^{|\sfb|} \Res(\sfb \beta) + (-1)^{|\sfc|} \Res(\sfc \gamma) .
\end{array}
\]
%\eeqn
%defined by
%\beqn
%(\beta, \gamma) \mapsto (-1)^{|\sfb|} \Res(\sfb \beta) + (-1)^{|\sfc|} \Res(\sfc \gamma) .
%\eeqn
Likewise, we can view a quadratic element in $\cP_2$ as a linear functional
\beqn
\Sym^2 \left(\laurent^{(2)} [1] \oplus \laurent \right) \to \C .
\eeqn
For example, the quadratic element $\sfb \sfc \in \cP_2$ corresponds to the quadratic functional
\beqn
\beta \gamma \mapsto (-1)^{|\sfb|+ |\sfc|} \Res (\sfb \beta) \Res(\sfc \gamma) ,
\eeqn
where $\beta \in \laurent^{(2)} [1]$ and $\gamma \in \laurent$.

Let $J$ be the map
\beqn
J \colon \laurent \to \cP_2
\eeqn
defined by sending an element $\alpha \in \sfA$ to the quadratic functional $J_\alpha \in \cP_2$ defined by
\beqn
J_\alpha (\beta \gamma) = \Res(\beta \alpha \gamma) .
\eeqn

\begin{lem}
Let $\alpha = z_1^m \lambda_2^n$.
Then $J_\alpha$ corresponds to the infinite series
\begin{multline}
J_\alpha = \sum_{a,b \geq 0} \sum_{k=0}^{a+b} c_k^{a,b} \sff^k(z_1^{a+m}\lambda_2^{b+n} \omega) \cdot \sff^{a+b-k} (z_1^b \lambda_2^a) \d^2 z \\
+ \sum_{a,b \geq 0} \sum_{k=0}^{a+b} c_k^{a,b} \sff^k(z_1^{a+m}\lambda_2^{b+n}) \cdot \sff^{a+b-k} (z_1^b \lambda_2^a \omega) \d^2 z ,
\end{multline}
for some constants $c_{k}^{a,b}$.

Similarly, let $\alpha = z_1^m \lambda_2^n \omega$.
Then $J_\alpha$ corresponds to the infinite series
\beqn
J_\alpha = \sum_{a,b \geq 0} \sum_{k=0}^{a+b} c_k^{a,b} \sff^k(z_1^{a+m}\lambda_2^{b+n} \omega) \cdot \sff^{a+b-k} (z_1^b \lambda_2^a \omega) \d^2 z ,
\eeqn
which is of degree $+1$.
\end{lem}

\begin{prop}
The following are true of the map $J$:
\begin{enumerate}
\item $J$ is a cochain map.
\item $J$ is $\lie{sl}(2)$-equivariant.
\item For $a,a' \in \laurent$ one has $\{J_a, J_{a'}\} = 0$.
\end{enumerate}
In particular, if we endow $\sfA$ with the structure of an abelian dg Lie algebra then $J$ is a map of dg Lie algebras.
\end{prop}
\begin{proof}
For the first claim observe that
\beqn
J_{\dbar a} (\beta \gamma) = \Res(\beta (\dbar a) \gamma) = \pm \Res((\dbar \beta) a \gamma) \pm \Res(\beta a \dbar \gamma) = J_a (\dbar \beta \gamma) + (-1)^{|\beta|} J_a(\beta \dbar \gamma) 
\eeqn
by (3) in lemma \ref{lem:res}.
\end{proof}

\section{Recollection on anomalies for quantum moment maps}

Let $V$ be a dg vector space and let $V^* = \Hom(V, \CC)$ denote its dual complex. 
Consider the dg Poisson algebra $\cP = \Sym(V \oplus V^*)$ whose Poisson bracket extends the skew-symmetrized evaluation pairing on $V \oplus V^*$.
We view $\cP$ as functions on the cotangent bundle~$T^* V$.
Observe that the quadratic functions form a dg Lie subalgebra
\[
\cP_2 = (\Sym^2(V \oplus V^*), \{-,-\})
\]
of $\cP$.
Note that for $V$ a finite-dimensional vector space concentrated in degree zero,
there is a natural identification $\cP_2 = \frak{sp}(T^*V)$, which plays an important role in corollary~\ref{cor:trivforsp}.

There is a natural deformation quantization of $\cP$ as a Weyl algebra $\cW$.
Observe that there is a dg Lie subalgebra
\[
\cW_2 = (\Sym^2(V \oplus V^*) \oplus \Sym^0(V \oplus V^*), [-,-])
\]
of $\cW$.
It is a central extension of $\cP_2$; let $\omega \in H^2(\cP_2)$ denote the associated 2-cocycle.

Suppose that there is a Lie algebra map $J: \fg \to \cP_2$.
We view as a special class of symmetries for the free theory encoded by $\cP$:
these are the symmetries where $\fg$ deforms the system to another {\em free} system.
This situation appears in the free field realization we are studying.

\begin{lem}
Given a Lie algebra map $J: \fg \to \cP_2$, the obstruction to $J$ being a map of Lie algebras $\fg \to \cW_2$ is the 2-cocycle $J^* \omega \in H^2(\fg)$.
\end{lem}

We call $J^*\omega$ the {\em anomaly} for quantizing this symmetry.

\begin{cor}
If the anomaly $J^*\omega$ does not vanish, then there is an $L_\infty$ algebra map $J^q: \widehat{\fg} \to \cW_2$ out of the $L_\infty$ extension
\[
\CC \to \widehat{\fg} \to \fg
\]
determined by $J^* \omega$.
\end{cor}

On the other hand, this phenomenon does not happen in the most traditional situations.

\begin{cor}
\label{cor:trivforsp}
For $V$ a finite-dimensional vector space concentrated in degree zero, the anomaly $\omega$ is trivial as $\frak{sp}(T^* V) = \cP_2$ is simple and hence $H^2(\cP_2) = 0$.
\end{cor}

In other words, quadratic Hamiltonians (elements of $\cP_2$) admit a natural quantization (lift to $\cW_2$).
Normal ordering offers a method, but there are other choices.

\owen{I am not sure what the status is for infinite-dimensional vector spaces. It seems like functional analysis plays a nontrivial role.}

\section{Recollections on one-dimensional algebraic free field realization}

Let $\CC^\infty$ denote the direct sum of countably many copies of $\CC$.
We will fix a basis $\{x_n\}_{n \in \ZZ}$.
This vector space can also be described as the colimit of finite-dimensional spaces under inclusion 
\[
\CC^{2N+1} \cong \bigoplus_{n = -N}^N \CC x_n \hookrightarrow \bigoplus_{n = -N-1}^{N+1} \CC x_n \cong \CC^{2N+3}.
\]
Later we will identify this vector space with Laurent polynomials $\CC[t,t^{-1}]$ by identifying $t^n$ with~$x_n$.

There are then two distinct infinite-dimensional Lie algebras that naturally act on this space:
\[
\gl_\infty \hookrightarrow \End(\CC^\infty).
\]
Here
\[
\gl_\infty = \colim \gl_{2N+1} = \bigoplus_{m,n \in \ZZ} \Hom(\CC x_m, \CC x_n),
\]
and, in concrete terms, it corresponds to those matrices acting on $\CC^\infty$ that have finitely many nonzero entries.
By contrast, 
\[
\End(\CC^\infty) = \Hom \left(\bigoplus_{m \in \ZZ} \CC x_m, \bigoplus_{n \in \ZZ} \CC x_n \right) \cong \prod_{m\in \ZZ} \bigoplus_{n \in \ZZ} \Hom(\CC x_m, \CC x_n),
\]
and, in concrete terms, it corresponds to those matrices where each column has finitely many nonzero entries but a row can have arbitrarily many nonzero entries.

In between these two sits an interesting Lie algebra $\Bar{\gl}_\infty$ consisting of matrices with finitely many nonzero {\em diagonals}.

This Lie algebra has two features crucial for us:
\begin{enumerate}
\item It contains the shift operators $S_k: x_n \mapsto x_{n+k}$ and hence contains $\CC[t,t^{-1}]$ as a subalgebra (note that multiplication by $t^k$ corresponds to $S_k$). More generally, every loop algebra $\gl_N(\CC[t,t^{-1}])$ embeds into~$\Bar{\gl}_\infty$.
\item It has a nontrivial 2-cocycle $\Bar{\alpha}$ where
\[
\Bar{\alpha}(E_{ij}, E_{mn}) =
\begin{cases}
1, & i = n, j = m, i \leq 0, j \geq 1\\
-1, & i = n, j = m, n \leq 0, m \geq 1\\
0, & \text{otherwise}
\end{cases}.
\]
\end{enumerate}
These features interact because for any embedding $\iota$ of a loop algebra, 
the pullback $\iota^*\Bar{\alpha}$ determines a nontrivial central extension: 
it yields the famous affine Lie algebra using the residue pairing.
(One can see $\Bar{\alpha}$ is nontrivial by using the fact that the affine Lie algebra is a nontrivial extension.)

\begin{rmk}
Kac and Peterson (as well as Date, Jimbo, et al) found this extension by $\Bar{\alpha}$ because they'd found that $\Bar{\gl}_\infty$ has a {\em projective} representation on the infinite wedge space.
By the embedding fact, {\em every} affine Lie algebra also acts on the infinite wedge space.
\end{rmk}

\section{Trying to mimic this strategy}

We will try to replicate the strategy above but replacing $\CC[t,t^{-1}]$ by $\sfH = H^\bullet(\sfA)$.
Recall that
\[
\sfH^\bullet = 
\begin{cases}
\CC[z_1,z_2],&\bullet = 0\\
\CC[z_1^{-1}, z_2^{-1}] \omega,&\bullet = 1\\
0,&\text{otherwise}\\
\end{cases}
\]
where $\omega$ denotes the Martinelli-Bochner form
\[
\frac{\zbar_2 \d\zbar_1 - \zbar_1 \d\zbar_2}{(2\pi i)^2|z_1 z_2|^2},
\]
which is the two-dimensional analog of $\d z/2\pi i z$.
Note that we have a natural basis for $\sfH$ via monomials $z_1^a z_2^b$, with $a,b \in \NN$, and $z_1^{-a} z_2^{-b} \omega$, with $a,b \in \NN$.

There is a natural {\em residue} operation on $\sfH$,
which simply takes the coefficient on $\omega$ and vanishes on a nontrivial monomial times~$\omega$.
In terms of calculus, it means that one wedges with $\d^2 z$ and then integrates over the unit ball.
This residue determines a pairing on $\sfH$ by mimicking the one-dimensional case:
\[
(z_1^m z_2^n, z_1^{-k} z_2^{-l} \omega) =  \delta_{mk} \delta_{nl},
\]
which has degree -1 and hence vanishes between elements of the same degree.

In a similar way, we introduce a trilinear form of degree -1 by
\[
\alpha_{(2)}( f, g, h \omega) = \int_{|z| \leq 1} \partial(f) \wedge \partial(g) \wedge h \omega.
\]
\owen{Up to a constant that I'll worry about later.}
In terms of monomials, it is
\[
\alpha_{(2)}( z_1^a z_2^b , z_1^m z_2^n, z_1^{-k} z_2^{-l} \omega) = (an - bm) \delta_{(a+m-1)k} \delta_{(b+n-1)l}.
\]
This formula actually determines a nontrivial 2-cocycle on $\sfH$, viewed as an abelian graded  Lie algebra.

We want to find a 2-cocycle $\Bar{\alpha}_{(2)}$ on a graded Lie algebra
\[
\Bar{\gl}_\infty^{(2)} \subset \End(\sfH)
\]
that pulls back to $\alpha_{(2)}$ along an embedding $\sfH \hookrightarrow \Bar{\gl}_\infty^{(2)}$ that factors the canonical inclusion $\sfH \hookrightarrow \End(\sfH)$ sending an element $x$ to left multiplication by~$x$.
This situation would be parallel to the one-dimensional construction.

\subsection{Defining the Lie algebra}

Here is a cheeky and inexplicit way to define a $\Bar{\gl}_\infty^{(2)}$: take the smallest sub-associative algebra of $\End(\sfH)$ that contains both $\sfH$ and $\gl_\infty^{(2)}$ (which is simply the direct sum of the homs between basis elements). 
\owen{I think this works but it's ``smaller'' than the corresponding one-dimensional $\Bar{\gl_\infty}$.
We could instead ask to include derivatives also, so that we contain differential operators.}

As an alternative, we characterize what we mean by ``diagonals,'' as follows.

We introduce the following notation for elements of $\gl_\infty^{(2)}$:
\begin{itemize}
\item Let $E^{00}_{ab, cd}$ denote the matrix sending $z_1^a z_2^b$ to $z_1^c z_2^d$.
\item Let $E^{11}_{ab, cd}$ denote the matrix sending $z_1^{-a} z_2^{-b} \omega$ to $z_1^{-c} z_2^{-d} \omega$.
\item Let $E^{01}_{ab, cd}$ denote the matrix sending $z_1^a z_2^b$ to $z_1^{-c} z_2^{-d} \omega$.
\item Let $E^{10}_{ab, cd}$ denote the matrix sending $z_1^{-a} z_2^{-b} \omega$ to $z_1^c z_2^d $.
\end{itemize}
Note that the superscript $ij$ indicates whether the matrix the degree of the domain and codomain, respectively.
The subscripts $a,b,c,d$ are all nonnegative integers.

\begin{dfn}
Given a matrix $M \in \End(\sfH)$, its {\em support} is the set of indices $(i, j; a, b, c, d)$ for which its coefficient $M_i$ are not zero.
\end{dfn}

For example, the support of $E^{00}_{ab, cd}$ is $(0,0,a, b,c, d)$, 
while the support of the identity is $\{(i,i;a,b,a,b) \,: \, i \in \{0,1\}, a,b \in \NN\}$.

Consider now the operators arising from multiplication.
In the one-dimensional case, the support of such multiplication operators (or shift operators) characterizes the diagonals,
so we examine the two-dimensional structure at play.
\begin{itemize}
\item Multiplying by $z_1^m z_2^n$ is a linear map sending $z_1^a z_2^b$ to $z_1^{m+a} z_2^{n+b}$ and sending $z_1^{-a} z_2^{-b} \omega$ to $z_1^{m-a} z_2^{n-b} \omega$ for $m -a \leq 0$ and $n-b \leq 0$ (otherwise, the element maps to zero).
Its support is 
\[
\{(0,0;a,b,m+a,n+b) \,: \, a,b \in \NN\} \cup \{(1,1;a,b,a-m,b-n) \,: \, 0 \leq m \leq a, 0 \leq n \leq b\}.
\]
\item Multiplying by $z_1^{-m} z_2^{-n} \omega$ annihilates anything in degree 1 and sends $z_1^a z_2^b$ to $z_1^{a-m} z_2^{b-n}$ for $a-m \leq 0$ and $b-n \leq 0$ (otherwise, the elements maps to zero). 
Its support is 
\[
\{(0,1;a,b,m-a,n-b) \,: \, 0 \leq a \leq m, 0 \leq b \leq n\}.
\]
\end{itemize}
In light of this support structure, we pose the following definition.

\begin{dfn}
A matrix $M in \End(\sfH)$ has {\em finitely many diagonals} if its support is equal to the union of finitely many multiplication operators.
\end{dfn}

One could then define $\Bar{\gl}_\infty^{(2)}$ as matrices with finitely many diagonals.

\subsection{Defining the cocycle}

Here is a guess for $\Bar{\alpha}_{(2)}$.
It vanishes unless there are two entries of degree 0 and one entry of degree 1.
We define
\[
\Bar{\alpha}(E^0_{ab}, E^0_{mn}, E^{+}_{kl}) = \delta_{(a+m)k} \delta_{(b+n)l}
\]
and impose the relevant sign for permuting the entries.
\owen{This description is under-determined. I need to improve it.}

\end{document}